Crosslinguistic Word Orders Enable and Efficient Tradeoff between Memory and Surprisal

### 2019-03-02 paper.tex
results
paper

\title{Word Orders Trade Off Predictability and Memory Load Across Languages}%  Word Order Minimizes  Across Languages
\author{Michael Hahn, Judith Degen, Richard Futrell}
\date{2018}

@@

Are languages optimized for processing with limited memory?
Memory limitations are well-established as a factor impacting sentence processing, and have been argued to account for crosslinguistic word order regularities.
Computational models of language processing implementing memory constraints take various forms and make different kinds of assumptions about the architecture underlying human language processing.
We first  establish general information-theoretic lower bounds on memory that hold in any model of language processing, applying both to speakers and listeners. %assuming only that listeners perform incremental prediction.
Applying these results to corpora from over 50 languages, we then provide evidence that word orders in human language are optimized for memory demands of speakers and listeners. % producing language and listeners predicting input.
\end{abstract}

@@

@@

\section{Results (Morphology on UD Treebanks)}\label{sec:results-table}

Statistics reported:

\begin{enumerate}
%	\item CIs for $C_x := B_x - A_x$ may be a resonable stopping criterion
	\item $A_x := P_0(\{X : X_1 \leq x_1 \wedge X_2 \leq x_2\})$
	\item $B_x := P_0(\{X : X_1 \geq x_1 \wedge X_2 \geq x_2\})$
	\item  $E := \frac{\E_{x \sim P_1}[A_x]}{\E_{x \sim P_1}[A_x+B_x]}$
	\item Compute $A_{x'}$ for each $x'$ in the baseline distribution, and then calculate the percentile rank of $A_x$ among these values:
$D := P_0(A_X \leq \E_{x \sim P_1}[A_x])$
\item For each sample $x$ from real orderings, we look at the proportion $W(x)$ of samples from the baseline languages that are below $x$ throughout the entire range where both curves are defined.

\end{enumerate}

For each language, we show the number of sentences in the training partition, the surprisal-speaker memory tradeoff plane, the listener's surprisal-memory tradeoff curves (smoothed means), and the statistics defined above with bootstraped CIs.

Languages where the stopping criterion hasn't been reached are marked with asterisks.

\begin{longtable}{lllllllllllllll}
Afrikaans  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Afrikaans-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Afrikaans-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.13  &  [0.05, 0.28]  \\ 
1315  &    &    &  $B_x-A_x$  &  0.0  &  [-0.0, 0.01]  \\ 
  &    &    &  $E_x$  &  0.42  &  [0.0, 1.0]  \\ 
  &    &    &  $W_x$  &  0.99  &  [0.96, 1.0]  \\ [10.25ex] \hline
Amharic-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Amharic-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Amharic-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.16  &  [0.06, 0.29]  \\ 
974  &    &    &  $B_x-A_x$  &  0.1  &  [0.04, 0.18]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Arabic*  &\\ [10.25ex] \hline
Armenian-Adap*  &\\ [10.25ex] \hline
Basque  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Basque-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Basque-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.09  &  [0.03, 0.15]  \\ 
5396  &    &    &  $B_x-A_x$  &  0.82  &  [0.76, 0.88]  \\ 
  &    &    &  $E_x$  &  0.99  &  [0.97, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Breton-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Breton-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Breton-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.11  &  [0.03, 0.22]  \\ 
788  &    &    &  $B_x-A_x$  &  0.32  &  [0.2, 0.42]  \\ 
  &    &    &  $E_x$  &  0.99  &  [0.95, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Bulgarian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Bulgarian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Bulgarian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.07  &  [0.0, 0.22]  \\ 
8907  &    &    &  $B_x-A_x$  &  0.56  &  [0.44, 0.7]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Buryat-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Buryat-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Buryat-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.09  &  [0.02, 0.19]  \\ 
808  &    &    &  $B_x-A_x$  &  0.28  &  [0.19, 0.42]  \\ 
  &    &    &  $E_x$  &  1.0  &  [0.99, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Cantonese-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Cantonese-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Cantonese-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.17  &  [0.08, 0.31]  \\ 
550  &    &    &  $B_x-A_x$  &  0.25  &  [0.17, 0.36]  \\ 
  &    &    &  $E_x$  &  0.92  &  [0.85, 0.98]  \\ 
  &    &    &  $W_x$  &  0.94  &  [0.86, 1.0]  \\ [10.25ex] \hline
Catalan  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Catalan-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Catalan-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.13  &  [0.03, 0.25]  \\ 
13123  &    &    &  $B_x-A_x$  &  0.44  &  [0.33, 0.57]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Chinese  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Chinese-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Chinese-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.12  &  [0.04, 0.23]  \\ 
3997  &    &    &  $B_x-A_x$  &  0.1  &  [0.05, 0.19]  \\ 
  &    &    &  $E_x$  &  0.99  &  [0.95, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Croatian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Croatian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Croatian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.1  &  [0.03, 0.23]  \\ 
7689  &    &    &  $B_x-A_x$  &  0.13  &  [0.05, 0.23]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Czech  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Czech-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Czech-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.14  &  [0.05, 0.27]  \\ 
102993  &    &    &  $B_x-A_x$  &  0.66  &  [0.56, 0.77]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  0.94  &  [0.86, 0.99]  \\ [10.25ex] \hline
Danish  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Danish-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Danish-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.11  &  [0.02, 0.26]  \\ 
4383  &    &    &  $B_x-A_x$  &  0.15  &  [0.06, 0.25]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Dutch  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Dutch-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Dutch-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.07  &  [0.01, 0.15]  \\ 
18310  &    &    &  $B_x-A_x$  &  0.04  &  [0.01, 0.08]  \\ 
  &    &    &  $E_x$  &  0.88  &  [0.73, 1.0]  \\ 
  &    &    &  $W_x$  &  0.98  &  [0.96, 1.0]  \\ [10.25ex] \hline
Estonian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Estonian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Estonian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.39  &  [0.3, 0.48]  \\ 
6959  &    &    &  $B_x-A_x$  &  0.26  &  [0.18, 0.36]  \\ 
  &    &    &  $E_x$  &  0.76  &  [0.67, 0.84]  \\ 
  &    &    &  $W_x$  &  0.72  &  [0.65, 0.79]  \\ [10.25ex] \hline
Faroese-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Faroese-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Faroese-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.13  &  [0.05, 0.26]  \\ 
1108  &    &    &  $B_x-A_x$  &  0.34  &  [0.24, 0.45]  \\ 
  &    &    &  $E_x$  &  1.0  &  [0.99, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Finnish  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Finnish-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Finnish-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.05  &  [0.01, 0.12]  \\ 
27198  &    &    &  $B_x-A_x$  &  0.95  &  [0.91, 0.99]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
French*  &\\ [10.25ex] \hline
German  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/German-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/German-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.44  &  [0.37, 0.51]  \\ 
13814  &    &    &  $B_x-A_x$  &  0.19  &  [0.12, 0.29]  \\ 
  &    &    &  $E_x$  &  0.66  &  [0.6, 0.74]  \\ 
  &    &    &  $W_x$  &  0.63  &  [0.57, 0.69]  \\ [10.25ex] \hline
Greek  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Greek-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Greek-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.07  &  [0.0, 0.2]  \\ 
1662  &    &    &  $B_x-A_x$  &  0.59  &  [0.48, 0.73]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Hebrew  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hebrew-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hebrew-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.12  &  [0.03, 0.24]  \\ 
5241  &    &    &  $B_x-A_x$  &  0.19  &  [0.08, 0.3]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Hindi  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hindi-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hindi-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.07  &  [0.0, 0.24]  \\ 
13304  &    &    &  $B_x-A_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Hungarian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hungarian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Hungarian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.55  &  [0.44, 0.65]  \\ 
910  &    &    &  $B_x-A_x$  &  0.07  &  [0.0, 0.15]  \\ 
  &    &    &  $E_x$  &  0.62  &  [0.51, 0.74]  \\ 
  &    &    &  $W_x$  &  0.83  &  [0.76, 0.89]  \\ [10.25ex] \hline
Indonesian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Indonesian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Indonesian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.36  &  [0.28, 0.43]  \\ 
4477  &    &    &  $B_x-A_x$  &  0.27  &  [0.22, 0.33]  \\ 
  &    &    &  $E_x$  &  0.82  &  [0.77, 0.87]  \\ 
  &    &    &  $W_x$  &  0.62  &  [0.57, 0.66]  \\ [10.25ex] \hline
Japanese  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Japanese-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Japanese-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.05  &  [0.0, 0.18]  \\ 
7164  &    &    &  $B_x-A_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Kazakh-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Kazakh-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Kazakh-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.08  &  [0.0, 0.24]  \\ 
947  &    &    &  $B_x-A_x$  &  0.29  &  [0.2, 0.38]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Kurmanji-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Kurmanji-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Kurmanji-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.21  &  [0.09, 0.36]  \\ 
634  &    &    &  $B_x-A_x$  &  0.1  &  [0.05, 0.16]  \\ 
  &    &    &  $E_x$  &  0.86  &  [0.72, 0.95]  \\ 
  &    &    &  $W_x$  &  0.96  &  [0.92, 1.0]  \\ [10.25ex] \hline
Latvian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Latvian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Latvian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.51  &  [0.41, 0.62]  \\ 
4124  &    &    &  $B_x-A_x$  &  -0.05  &  [-0.11, 0.01]  \\ 
  &    &    &  $E_x$  &  0.37  &  [0.24, 0.53]  \\ 
  &    &    &  $W_x$  &  0.66  &  [0.58, 0.74]  \\ [10.25ex] \hline
Naija-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Naija-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Naija-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.1  &  [0.02, 0.25]  \\ 
848  &    &    &  $B_x-A_x$  &  0.33  &  [0.21, 0.47]  \\ 
  &    &    &  $E_x$  &  1.0  &  [0.99, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
North Sami  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/North_Sami-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/North_Sami-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.36  &  [0.28, 0.45]  \\ 
2257  &    &    &  $B_x-A_x$  &  0.15  &  [0.08, 0.23]  \\ 
  &    &    &  $E_x$  &  0.73  &  [0.62, 0.81]  \\ 
  &    &    &  $W_x$  &  0.36  &  [0.28, 0.43]  \\ [10.25ex] \hline
Norwegian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Norwegian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Norwegian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.14  &  [0.04, 0.3]  \\ 
29870  &    &    &  $B_x-A_x$  &  0.63  &  [0.48, 0.78]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Persian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Persian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Persian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.17  &  [0.07, 0.32]  \\ 
4798  &    &    &  $B_x-A_x$  &  0.12  &  [0.06, 0.2]  \\ 
  &    &    &  $E_x$  &  0.93  &  [0.77, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [0.99, 1.0]  \\ [10.25ex] \hline
Polish  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Polish-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Polish-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.39  &  [0.3, 0.5]  \\ 
6100  &    &    &  $B_x-A_x$  &  0.08  &  [-0.0, 0.15]  \\ 
  &    &    &  $E_x$  &  0.6  &  [0.5, 0.69]  \\ 
  &    &    &  $W_x$  &  0.31  &  [0.23, 0.39]  \\ [10.25ex] \hline
Romanian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Romanian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Romanian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.12  &  [0.03, 0.24]  \\ 
8664  &    &    &  $B_x-A_x$  &  0.15  &  [0.07, 0.24]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Serbian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Serbian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Serbian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.06  &  [0.01, 0.15]  \\ 
2935  &    &    &  $B_x-A_x$  &  0.19  &  [0.13, 0.26]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Slovak  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Slovak-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Slovak-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.21  &  [0.13, 0.3]  \\ 
8483  &    &    &  $B_x-A_x$  &  0.48  &  [0.39, 0.56]  \\ 
  &    &    &  $E_x$  &  0.94  &  [0.9, 0.97]  \\ 
  &    &    &  $W_x$  &  0.89  &  [0.83, 0.94]  \\ [10.25ex] \hline
Slovenian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Slovenian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Slovenian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.31  &  [0.22, 0.39]  \\ 
7532  &    &    &  $B_x-A_x$  &  0.44  &  [0.35, 0.56]  \\ 
  &    &    &  $E_x$  &  0.85  &  [0.79, 0.92]  \\ 
  &    &    &  $W_x$  &  0.79  &  [0.72, 0.85]  \\ [10.25ex] \hline
Spanish* &\\ [10.25ex] \hline
Swedish  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Swedish-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Swedish-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.11  &  [0.02, 0.29]  \\ 
7041  &    &    &  $B_x-A_x$  &  0.06  &  [0.01, 0.12]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Tamil  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Tamil-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Tamil-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.15  &  [0.04, 0.29]  \\ 
400  &    &    &  $B_x-A_x$  &  0.17  &  [0.1, 0.28]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Thai-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Thai-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Thai-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.15  &  [0.06, 0.28]  \\ 
900  &    &    &  $B_x-A_x$  &  0.19  &  [0.12, 0.27]  \\ 
  &    &    &  $E_x$  &  0.99  &  [0.97, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Turkish  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Turkish-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Turkish-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.18  &  [0.05, 0.35]  \\ 
3685  &    &    &  $B_x-A_x$  &  0.03  &  [0.01, 0.06]  \\ 
  &    &    &  $E_x$  &  1.0  &  [1.0, 1.0]  \\ 
  &    &    &  $W_x$  &  1.0  &  [1.0, 1.0]  \\ [10.25ex] \hline
Ukrainian  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Ukrainian-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Ukrainian-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.12  &  [0.03, 0.22]  \\ 
4506  &    &    &  $B_x-A_x$  &  0.18  &  [0.1, 0.24]  \\ 
  &    &    &  $E_x$  &  0.94  &  [0.87, 1.0]  \\ 
  &    &    &  $W_x$  &  0.99  &  [0.97, 1.0]  \\ [10.25ex] \hline
Uyghur-Adap  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Uyghur-Adap-entropy-memory.pdf}}  &  \multirow{4}{*}{\includegraphics[width=0.25\textwidth]{figures/Uyghur-Adap-listener-surprisal-memory.pdf}}  &  $D_x$  &  0.17  &  [0.06, 0.3]  \\ 
1656  &    &    &  $B_x-A_x$  &  0.0  &  [-0.0, 0.01]  \\ 
  &    &    &  $E_x$  &  0.9  &  [0.0, 1.0]  \\ 
  &    &    &  $W_x$  &  0.0  &  [0.0, 0.0]  \\ [10.25ex] \hline
Vietnamese* &\\ [10.25ex] \hline
\end{longtable}

>>>>

\title{Crosslinguistic Word Orders Enable and Efficient Tradeoff between Memory and Surprisal}
\author{Michael Hahn, Judith Degen, Richard Futrell}
\date{2018}

@@

Online memory limitations are well-established as a factor impacting sentence processing and have been argued to account for crosslinguistic word order regularities. Building off expectation-based models of language processing, we provide an information-theoretic formalization of these memory limitations. We introduce the idea of a memory-surprisal tradeoff: comprehenders can achieve lower average surprisal per word at the cost of storing more information about past context. We show that the shape of the tradeoff is determined in part by word order. In particular, languages will enable more efficient tradeoffs when they exhibit information locality: when predictive information about a word is concentrated in the word’s recent past. We show evidence from corpora of 52 real languages showing that languages allow for more efficient memory-surprisal tradeoffs than random baseline word order grammars.

%Are languages optimized for processing with limited memory?
%Memory limitations are well-established as a factor impacting sentence processing, and have been argued to account for crosslinguistic word order regularities.
%Computational models of language processing implementing memory constraints take various forms and make different kinds of assumptions about the architecture underlying human language processing.
%We first  establish general information-theoretic lower bounds on memory that hold in any model of language processing, applying both to speakers and listeners. %assuming only that listeners perform incremental prediction.
%Applying these results to corpora from over 50 languages, we then provide evidence that word orders in human language are optimized for memory demands of speakers and listeners. % producing language and listeners predicting input.
\end{abstract}

@@

\begin{longtable}{lllllllllllllll}
\input{results-table-word-level.tex}
\end{longtable}

@@
